---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# geodk

<!-- badges: start -->
[![R-CMD-check](https://github.com/aleksanderbl29/geodk/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/aleksanderbl29/geodk/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/aleksanderbl29/geodk/graph/badge.svg)](https://app.codecov.io/gh/aleksanderbl29/geodk)
[![Project Status: WIP – Development in progress](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
<!-- badges: end -->

`{geodk}` provides access to Danish geospatial data.

## Installation

You can install the development version of geodk from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("aleksanderbl29/geodk")
```



## Installation

```{r check_cran, include=FALSE}
if (!require(available)) install.packages("available")
is_on_cran <- !available::available_on_cran("geodk")
```

```{r cran_instruct, echo=FALSE, results='asis', eval=is_on_cran}
cat("You can install `{geodk}` from CRAN with:")
```

```{r cran_code, eval=FALSE, include=is_on_cran}
install.packages("geodk")
```

```{r r_universe_instruct, echo=FALSE, results='asis'}
if (is_on_cran) {
  cat("Or you can install the latest pre-release version of `{geodk}` from r-universe with:")
} else if (!is_on_cran) {
  cat("You can install `{geodk}` from r-universe with:")
}
```

```{r r_universe, eval=FALSE}
install.packages(
  "geodk",
  repos = c(
    aleksanderbl29 = "https://aleksanderbl29.r-universe.dev",
    getOption("repos")
  )
)
```

You can install the latest development version of `{geodk}` from [GitHub](https://github.com/aleksanderbl29/geodk) with:

``` r
# install.packages("devtools")
devtools::install_github("aleksanderbl29/geodk")
```

## Functions

`{geodk}` provides to main sets of functions:

- Data retrieval
- Plotting

### Plotting

The plotting functions include `plot_denmark()` which easily plots a basic map of Denmark and returns it as a `{ggplot2}` object for you to edit. It also includes `plot_region()` and `plot_municipality()` which plots a subset of Denmark with focus on provided regions or municipalities.

```{r example}
library(geodk)
plot_denmark()
```

```{r}
region <- plot_regions(region = c("Region Nordjylland", "Region Midtjylland"))
```

```{r}
municipality <- plot_municipalities(municipality = c("Aarhus", "Favrskov", "Vejle"))
```

```{r}
library(patchwork)
region + municipality
```
