---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# geodk

<!-- badges: start -->
[![R-CMD-check](https://github.com/aleksanderbl29/geodk/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/aleksanderbl29/geodk/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/aleksanderbl29/geodk/graph/badge.svg)](https://app.codecov.io/gh/aleksanderbl29/geodk)
<!-- badges: end -->

`{geodk}` provides access to Danish geospatial data.

## Installation

You can install the development version of geodk from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("aleksanderbl29/geodk")
```

## Functions

`{geodk}` provides to main sets of functions:

- Data retrieval
- Plotting

### Plotting

The plotting functions include `plot_denmark()` which easily plots a basic map of Denmark and returns it as a `{ggplot2}` object for you to edit. It also includes `plot_region()` and `plot_municipality()` which plots a subset of Denmark with focus on provided regions or municipalities.

```{r example}
library(geodk)
plot_denmark()
```

```{r}
region <- plot_regions(region = c("Region Nordjylland", "Region Midtjylland"))
```

```{r}
municipality <- plot_municipalities(municipality = c("Aarhus", "Favrskov", "Vejle"))
```

```{r}
library(patchwork)
region + municipality
```
